\documentclass[letterpaper,12pt]{article}
\input{preamble}

\chead{Transpose of a Matrix}

\begin{document}

\section*{Transpose of a Matrix}
\begin{definition}
Let $A$ be a $m \times n$ matrix. The \textbf{transpose} of $A$, $A^T$, is the matrix that results from switching the rows and columns of $A$. The $j$th column of $A^T$ is the $j$th row of $A$, and vise versa.
\end{definition}

\begin{example}
If $A = \begin{bmatrix} a & b \\ c & d \end{bmatrix}$, then $A^T = \begin{bmatrix} a & c \\ b & d \end{bmatrix}$.
\end{example}

\section*{Transpose of a Matrix}
\begin{definition}
Let $A$ be a $m \times n$ matrix. The \textbf{transpose} of $A$, $A^T$, is the matrix that results from switching the rows and columns of $A$. The $k$th row becomes the $k$th column, and vise versa. In other words, if $A = \begin{bmatrix} a_{ij} \end{bmatrix}$, then
\begin{align*}
    (A^T)_{ij} = a_{ji}
\end{align*}
In particular, the $(i,j)$-entry of $A^T$ is the $(j,i)$-entry of $A$.
\end{definition}

Note that transposition has higher precedence of matrix multiplication and addition, so $AB^T$ represents $A(B^T)$, and $A + B^T = A + (B^T)$, as opposed to $(AB)^T$ and $(A + B)^T$ which is the transpose of the product matrix $AB$ and sum $A + B$, respectively.
\\ \\ Intuitively, transposition is a kind of ``reflection".

\section*{Basic Properties of the Transpose}
\begin{theorem}
Let $A, B$ be $m \times n$ matrices, $c \in \mathbb{R}$. Then,
\begin{enumerate}[(a)]
    \item \textbf{Transpose of a transpose}.
    \begin{equation*}
        (A^T)^T = A
    \end{equation*}
    \item \textbf{Transpose of a sum}.
    \begin{equation*}
        (A + B)^T = A^T + B^T
    \end{equation*}
    \item \textbf{Transpose of a scalar multiple}.
    \begin{equation*}
        (cA)^T = cA^T
    \end{equation*}
\end{enumerate}
\end{theorem}

In words, (a) says that the transpose of a transpose is equal to the original matrix. Properties (b) and (c) together mean that taking the transpose is a ``linear" operation. These properties are intuitively true, however the proofs are quite dense with subtle notation and subscripts.

\begin{proof}
Let $A = \begin{bmatrix} a_{ij} \end{bmatrix}, B = \begin{bmatrix} b_{ij} \end{bmatrix}$.
\begin{enumerate}[(a)]
    \item The $(i,j)$-entry of $A^T$ is the $(j,i)$-entry of $A$. Then, the $(i,j)$-entry of $(A^T)^T$ is the $(j,i)$-entry of $A^T$, which in turn is the $(i,j)$-entry of $A$. Thus, $(A^T)^T = A$. In notation, $(A^T)_{ij} = a_{ji}$, so $(A^T)^T_{ij} = a_{ij} = A_{ij}$, and so $(A^T)^T = A$.
    \item Then, $(A^T)_{ij} = a_{ji}$ and $(B^T)_{ij} = b_{ji}$. Also, $(A + B)_{ij} = a_{ij} + b_{ij}$. Then,
    \begin{align*}
        \underbrace{(A + B)^T_{ij}}_{\text{$(i,j)$-entry of $(A+B)^T$}} & = (A + B)_{ji} && \text{by definition of transpose} \\
        & = a_{ji} + b_{ji} && \text{as the $(j,i)$-entry of $A + B$ is $a_{ji} + b_{ji}$} \\
        & = (A^T)_{ij} + (B^T)_{ij}
    \end{align*}
    \item The $(i,j)$-entry of $cA$ is $ca_{ij}$ or $(cA)_{ij} = ca_{ij}$. Then, the $(i,j)$-entry of $(cA)^T$ is the $(j,i)$-entry of $cA$ which is $ca_{ji}$, and so $((cA)^T)_{ij} = ca_{ji}$. Thus, $(cA)^T = cA^T$.
\end{enumerate}
\end{proof}

\section*{Transpose of a Product}
\begin{theorem}
Let $A$ be an $m \times n$ matrix, $B$ be an $n \times p$ matrix. Then, the transpose of the product $BA$ is the product of the transposes in reverse order, or
\begin{equation*}
    \boxed{(AB)^T = B^T A^T}
\end{equation*}
\end{theorem}

In particular, $(AB)^T \neq A^T B^T$, even if the dimensions of $A^T, B^T$ are such that the product $A^T B^T$ is defined.

\begin{proof}
\begin{align*}
    (AB)^t_{ij} & = (AB)_{ji} && \text{definition of transpose} \\
    & = \sum_{k=1}^n A_{jk} B_{ki} && \text{definition of matrix multiplication} \\
    & = \sum_{k=1}^n A^t_{kj} B^t_{ik} && \text{definition of transpose} \\
    & = \sum_{k=1}^n B^t_{ik} A^t_{kj} \\
    (AB)^t_{ij} & = (B^t A^t)_{ij} && \text{definition of matrix multiplication}
\end{align*}
\end{proof}

More generally, the transpose of a product of matrices is equal to the product of their transposes in \textit{reverse order},
\begin{equation*}
    \boxed{(A_1 A_2 \cdots A_n)^T = A_n^T \cdots A_2^T A_1^T}
\end{equation*}

\end{document}