\documentclass[letterpaper,12pt]{article}
\input{preamble}

\chead{Solution Sets of Linear Equations}

\begin{document}

\section*{Homogeneous Systems}
\begin{definition}
A system of linear equations is \textbf{homogeneous} if it can be written in the form $A\vec{x} = \vec{0}$, where $A$ is an $m \times n$ matrix and $\vec{0}$ is the zero vector in $\mathbb{R}^m$.
\begin{itemize}
    \item Otherwise, a system $A\vec{x} = \vec{b}$ is said to be \textbf{non-homogeneous}.
\end{itemize}
\end{definition}

Homogeneous systems always have at least one solution, in particular the solution $\vec{x} = \vec{0}$, often called the \textbf{trivial} solution. Then, the important question is whether the equation $A\vec{x} = \vec{0}$ has any \textbf{non-trivial} solutions, that is, solutions $\vec{x}$ which are not the zero vector.
\\ \\ Recall that if a system is consistent, then it has a unique solution if the REF of its augmented matrix has no free variables, and infinitely many solutions if there is at least one free variable. Then, for homogeneous systems,

\begin{theorem}
The homogeneous equation $A\vec{x} = \vec{0}$ has a non-trivial solution if and only if the equation has at least one free variable.
\end{theorem}

The solution set of a homogeneous equation can always be written as the span of some collection of vectors.
\begin{itemize}
    \item If the system has no non-trivial solutions, then the solution set is $\Span{\vec{0}} = \vec{0}$.
\end{itemize}

\section*{Nonhomogeneous Systems}
\begin{theorem}
Let $A\vec{x} = \vec{b}$ be a linear system, which is consistent with solution $\vec{v}_p$. Then, the solution set of $A\vec{x} = \vec{b}$ is the set of all vectors of the form $\vec{v} = \vec{v}_p + \vec{v}_h$, where $\vec{v}_h$ is any solution of the associated homogeneous system $A\vec{x} = \vec{0}$.
\end{theorem}



\section*{Solving a Consistent System in Parametric Vector Form}
\begin{enumerate}
    \item Use row reduction to convert the augmented matrix to RREF.
    \item Solve for each basic variable in terms of any free variables.
    \item Write a typical solution $\vec{x}$ as a vector which depends on any free variables.
    \item Decompose $\vec{x}$ into a linear combination of vectors with only numerical entries, and use free variables as parameters.
\end{enumerate}










\end{document}