\documentclass[letterpaper,12pt]{article}
\input{preamble}

\chead{Introduction to Matrices}

\begin{document}

\section*{Misc}


Matrices provide an efficient way to represent and solve systems of linear equations, and also to record data.

\section*{Matrices}
A matrix is an array of numbers arranged in (horizontal) rows and (vertical) columns, and within brackets.

\begin{definition}
Let $m, n \in \mathbb{N}$. An $m \times n$ \textbf{matrix} (read as ``$m$ by $n$ matrix"), is a rectangular array of $m$ rows and $n$ columns of real numbers,
\begin{equation*}
    \begin{bmatrix}
    a_{11} & a_{12} & \dots & a_{1n} \\
    a_{21} & a_{22} & & \vdots \\
    \vdots & & \ddots & \\
    a_{m1} & \dots & & a_{mn}
    \end{bmatrix}
\end{equation*}
Each \textbf{entry} (or \textbf{coefficient}, or \textbf{element}) in row $i$ and column $j$ is $a_{ij}$.
\begin{itemize}
    \item The plural of matrix is \textbf{matrices}.
    \item The \textbf{ith row} of a matrix $A$, is the row vector (or row matrix) $\begin{bmatrix} a_{i1} & \dots & a_{in} \end{bmatrix}$
    \item The \textbf{jth column} of a matrix $A$, is the column vector (or column matrix) $\begin{bmatrix} a_{1j} \\ \vdots \\ a_{mj} \end{bmatrix}$
    \item The set of all $m \times n$ matrices with coefficients in $\mathbb{F}$ is $M(m \times n, \mathbb{F})$
\end{itemize}
\end{definition}

Vectors can be represented as matrices. A row vector is a $1 \times n$ matrix (with 1 row and $n$ ``columns"), and a column vector is a $m \times 1$ matrix (with 1 column and $m$ ``rows").


\section*{Classification of Matrices}


\begin{example}
For example,
\begin{equation*}
    \begin{bmatrix} -4 & 4 \\ -5 & 9 \end{bmatrix}
\end{equation*}
is a $2 \times 2$ square matrix.
\end{example}


\section*{Upper/Lower Triangular Matrices}
\begin{definition}
$A$ is \textbf{upper triangular} if all the entries below the main diagonal are zero. In other words, for all $i > j$, $A_{ij} = 0$.
\end{definition}

\begin{definition}
$A$ is \textbf{lower triangular} if all the entries above the main diagonal are zero. In other words, for all $i < j$, $A_{ij} = 0$.
\end{definition}

\subsection*{Diagonal Matrices}
\begin{definition}
$A \in M(n \times n, \mathbb{F})$ is \textbf{diagonal} if and only if for all $i \neq j$, $A_{ij} = 0$. In other words, $A$ only has non-zero entries on the diagonal from top-left to bottom-right.
\end{definition}

\subsection*{Trace of a Matrix}
\begin{definition}
The \textbf{trace} of a matrix $A \in M(n \times n, \mathbb{F})$, $\operatorname{tr}(A)$, is the sum of the diagonal elements of $A$. In other words,
\begin{equation*}
    \operatorname{tr}(A) = A_{11} + A_{22} + \dots + A_{nn}
\end{equation*}
\end{definition}

\subsection*{Skew-Symmetric Matrices}
\begin{definition}
$A \in M(n \times n, \mathbb{F})$ is \textbf{skew-symmetric} if $A^T = -A$
\end{definition}

\section*{Matrix Operations Using a Graphing Calculator}




\end{document}